Adaptive Approximation for Multivariate Linear Problems \\ with Inputs Lying in a Cone

### 2018-11-02 RICAM2018NovTalk.tex

\newcommand{\sol}{S}
\newcommand{\app}{A}
%\DeclareMathOperator{\sol}{sol}
%\DeclareMathOperator{\app}{app}

@@

\iffalse
Adaptive Approximation to Multivariate Linear Problems for Inputs Lying in a Cone

@@

\renewcommand{\OffTitleLength}{-10ex}
\setlength{\FJHThankYouMessageOffset}{-8ex}
\title{Adaptive Approximation for Multivariate Linear Problems  \\ for Inputs Lying in a Cone}
\author[]{Fred J. Hickernell}
\institute{Department of Applied Mathematics \\
	Center for Interdisciplinary Scientific Computation \\  Illinois Institute of Technology \\
	\href{mailto:hickernell@iit.edu}{\url{hickernell@iit.edu}} \quad
	\href{http://mypages.iit.edu/~hickernell}{\url{mypages.iit.edu/~hickernell}}}

@@

	\newcommand{\smallscoop}{\parbox{1cm}{\includegraphics[width=1cm]{ProgramsImages/IceCreamScoop.eps}}\xspace}
	\newcommand{\medscoop}{\parbox{1.8cm}{\includegraphics[width=1.8cm]{ProgramsImages/IceCreamScoop.eps}}\xspace}
	\newcommand{\meddanger}{\parbox{0.8cm}{\vspace{-0.3cm}\includegraphics[width=0.75cm]{ProgramsImages/dangersign.eps}}\xspace}
	\newcommand{\medcone}{\parbox{1.2cm}{\includegraphics[width=0.55cm,angle=270]{ProgramsImages/MediumWaffleCone.eps}}\xspace}
	\newcommand{\largercone}{\parbox{2.2cm}{\vspace*{-0.2cm}\includegraphics[width=1cm,angle=270]{ProgramsImages/MediumWaffleCone.eps}}\xspace}
	\newcommand{\largecone}{\parbox{1.54cm}{\vspace*{-0.2cm}\includegraphics[width=0.7cm,angle=270]{ProgramsImages/MediumWaffleCone.eps}}\xspace}
	\newcommand{\smallcone}{\parbox{0.65cm}{\includegraphics[width=0.3cm,angle=270]{ProgramsImages/MediumWaffleCone.eps}}\xspace}

@@

\begin{frame}{Multivariate Linear Problems}
\vspace{-3ex}
\begin{tabular}{p{0.47\textwidth}p{0.5\textwidth}}
Given $f \in \cf$ find $g \in \cg$, where $g = \sol(f)$, \\
$\sol: \cf \to \cg$ is linear, e.g., 
\begin{gather*}
    g = \int_{\reals^d} f(\vx) \, \varrho(\vx) \, \dif \vx\\
    g = f\\
    g = \frac{\partial f}{\partial x_1} \\
    - \nabla^2 g = f, \ \  g = 0 \text{ on boundary}
\end{gather*}
&
\vspace{-10ex}
\begin{multline*}
    \ca(\ch) : = \{\app: \ch \times (0,\infty) \to \reals \text{ such that } \\
\norm[\cg]{\sol(f) - \app(f,\varepsilon) } \le \varepsilon \\ \forall f \in \ch \subseteq \cf, \ \varepsilon > 0 \}
\end{multline*}

\vspace{-2ex}
where $\app(f,\varepsilon)$ depends on \alert{function values}, \alert{Fourier coefficients}, or \alert{linear functionals}, e.g.,

\vspace{-4ex}
\begin{gather*}
    \app(f,\varepsilon) = \sum_{i=1}^n f(\vx_i) v_i, \quad v_i \in \cg \\
    \app(f,\varepsilon) = \sum_{i=1}^n \hf_i v_i, \quad v_i \in \cg \\
    \app(f,\varepsilon) = \sum_{i=1}^n L_i(f) v_i, \quad v_i \in \cg \\
\end{gather*}

\vspace{-4ex}
How big should $n$ be to satisfy error tolerance?

@@

\begin{tabular}{p{0.48\textwidth}p{0.49\textwidth}}
Given $f \in \cf$ find $g \in \cg$, where $g = \sol(f)$, \\
$\sol: \cf \to \cg$ is linear

\begin{multline*}
    \ca(\ch) : = \{\app: \ch \times (0,\infty) \to \reals \text{ such that } \\
\norm[\cg]{\sol(f) - \app(f,\varepsilon) } \le \varepsilon \\ \forall f \in \ch \subseteq \cf, \ \varepsilon > 0 \}
\end{multline*}

\vspace{-2ex}
where $\app(f,\varepsilon)$ depends on \alert{function values}, \alert{Fourier coefficients}, or \alert{linear functionals}
&

\vspace{-9ex}
\uncover<2->{\alert{Solvability}\footfullcite{KunEtal19a}: $\ca(\ch) \ne \emptyset$

@@

\alert{Construction}: Identify concrete $\app \in \ca(\ch)$}

@@

\uncover<3->{\alert{Cost}: $\cost(\app,f,\varepsilon) = $ \# of function data
\newline
 $\cost(\app,\ch,\varepsilon, \rho) = \max_{f \in \ch \cap \cb_{\rho}} \cost(\app,f,\varepsilon)$
 \newline
 $\cb_{\rho} = \{f \in \cf : \norm[\cf]{f} \le \rho \}$

@@

\alert{Complexity}\footfullcite{TraWasWoz88}: $\comp(\ca(\ch),\varepsilon,\rho)$ 
\newline \phantom{a} \hfill \hfill $= \min_{\app \in \ca(\ch)} \cost(\app,\ch,\varepsilon, \rho)$

@@

\alert{Optimality}:  \newline \phantom{a} \hfill \hfill $\cost(\app,\ch,\varepsilon, \rho) \le \comp(\ca(\ch),\alert{\omega} \varepsilon,\rho)$}

@@

\uncover<4->{\alert{Tractability}\footfullcite{NovWoz08a}: $\comp(\app,\ch,\varepsilon, \rho) \le C \rho^p\varepsilon^{-p} d^{q} $}

@@

@@

@@

\begin{frame}
\frametitle{What Does It Mean to Solve a Mathematics or Statistics Problem?}
\vspace{-5ex}

\[
\begin{array}{r@{\qquad}l}
	\text{Problem} & \text{Solution} \\
	\toprule
	x^2 +2 x - 3 = 0 &  x = -3, 1  \qquad \text{\alert{numbers}}\\
	\uncover<2->{\frac{\dif}{\dif x} \sin(x) & \cos(x) \qquad \text{\alert{elementary function}}} \\[2ex]
	\uncover<3->{\frac{1}{\sqrt{2\pi}} \int_{-\infty}^{1} \me^{-x^2/2} \, \dif x &  \Phi(1) \qquad \text{\alert{special function}} \\
	& \Phi \text{ is the standard Gaussian cumulative distribution function} } \\
\midrule
\uncover<4->{
\int_0^1 f(x) \, \dif x  & \text{Need a \alert{numerical solution}, i.e., } A(f,\varepsilon) \\[2ex]
f \text{ known via an algorithm} & \text{depending on integrand values } f(x_1), \, f(x_2), \ldots  \\
\text{i.e., can get } f(x) \ \forall x \in [0,1] & \text{such that } \abs{\int_0^1 f(x) \, \dif x - A(f,\varepsilon)} \le \varepsilon }
	\end{array}
	\]
\end{frame}

\begin{frame}
\frametitle{What About the Trapezoidal Rule?}
\vspace{-5ex}
\[
\int_0^1 f(x) \, \dif x \approx T_n(f)=\frac{1}{n}\left[\frac{f(0)}{2} + f\left(\frac{1}{n}\right) + \cdots + f\left(\frac{n-1}{n}\right) +  \frac{f(1)}{2} \right]
\]
\begin{tabular}{>{\raggedright}m{0.6\textwidth}@{\qquad}>{\centering}m{0.37\textwidth}}
\only<5->{\only<5-6>{Theory says
	\vspace{-3ex}
	\begin{equation*}
	\biggabs{\int_0^1 f(x) \, \dif x -T_n(f)} \le \frac{\overbrace{\int_0^1 \abs{f''(x)} \, \dif x}^{\text{curvature}}}{8 n^2} = \frac{\norm[1]{f''}}{8 n^2} \overset{\alert{?}}{\le} \varepsilon
	\end{equation*}

		\vspace{-1ex}
	\only<6>{Let $\cf = \{f :\norm[1]{f''} \le \rho \}$  \medscoop, and let}}\only<7->{Let}\uncover<6->{

	\vspace{-1ex}
	\[
	A(f,\varepsilon) = T_n(f) \quad \text{where }n = \left \lceil \sqrt{\frac{\rho}{8\varepsilon}} \, \right \rceil
	\]

	\vspace{-2ex}
	Then 
	\[
	 \abs{\int_0^1 f(x) \, \dif x - A(f,\varepsilon)} \le \varepsilon \quad \forall f \in \cf\only<7->{= \{f :\norm[1]{f''} \le \rho \}}, \, \varepsilon > 0
	 \]}}
	& 
	\only<1>{\includegraphics[width=5cm]{ProgramsImages/traprule-1.eps}\\$T_1(f)=0.5034$}
	\only<2>{\includegraphics[width=5cm]{ProgramsImages/traprule-2.eps}\\$T_2(f)=0.3949$}
	\only<3>{\includegraphics[width=5cm]{ProgramsImages/traprule-3.eps}\\$T_4(f)=0.3954$}
	\only<4->{\includegraphics[width=5cm]{ProgramsImages/traprule-4.eps}\\$T_{50}(f)=0.3957$}
\end{tabular}
\only<7->{We say that integration is \alert{solvable} for integrands in $\cf$  \medscoop}

\section{Solvability}
\begin{frame}
\frametitle{Solvability\footfullcite{KunEtal19a}}
\vspace{-4ex}

A problem is called \alert{solvable} for a set of inputs $\cf$ \alert{if there exists} a numerical algorithm that provides a solution with error no greater than $\varepsilon$.  Specifically, integration is solvable for $\cf$ if there exists an $A$ for which 
	\[
\abs{\int_0^1 f(x) \, \dif x - A(f,\varepsilon)} \le \varepsilon \quad \forall f \in \cf, \, \varepsilon > 0
\]
\uncover<2->{We demonstrated that integration is solvable for 
\[
\cf = \{f :\norm[1]{f''} \le \rho \}  \quad \medscoop
\]
\uncover<3>{Is integration solvable for 
\[
\cf = \{f :\norm[1]{f''} < \alert{\infty} \} \quad \text{since } \biggabs{\int_0^1 f(x) \, \dif x -T_n(f)} \le  \frac{\norm[1]{f''}}{8 n^2} \alert{?}
\]
}}

\begin{frame}
\frametitle{Integration Is \emph{Not} Solvable for $\cf = \{f :\norm[1]{f''} < \infty \}$}
\vspace{-3ex}
Suppose that there exists  $A$ such that 
\[
\abs{\int_0^1 f(x) \, \dif x - A(f,\varepsilon)} \le \varepsilon \quad \forall f \in \cf= \{f :\norm[1]{f''} < \infty \}, \, \varepsilon > 0
\]
\begin{tabular}{>{\raggedright}m{0.6\textwidth}@{\qquad}>{\centering}m{0.37\textwidth}}
	\begin{itemize}
	\item<2->Observe where $A(0,\varepsilon)$ gets function values

\item<3->Construct $f_{\text{bump}} \in \cf$, as tall as you like; note that $A(0,\varepsilon) = A(f_{\text{bump}},\varepsilon)$

	\item<4->  Since $\int_0^1 0 \, \dif x$ and $\int_0^1 f_{\text{bump}}(x) \, \dif x$ are so different, $A$ cannot be successful

\end{itemize}

	&
	\only<2>{\includegraphics[width=5cm]{ProgramsImages/ZeroFun.eps}}
	\only<3->{\includegraphics[width=5cm]{ProgramsImages/QuadHump.eps}}
\end{tabular}

\uncover<5->{Why doesn't this argument hold for $\cf = \{f :\norm[1]{f''} \le \rho \}  \medscoop$? \\  \uncover<6->{Because $f_{\text{bump}} \notin \medscoop$}}

\section{Optimality}

\begin{frame}
\frametitle{Is Trapezoidal Rule the Best Algorithm?}
\vspace{-4ex}
Let $\cost(A,f,\varepsilon)$ denote the number of function values, $n$ required for $A(f,\varepsilon)$ and 
\[
\cost(A,\cf,\varepsilon) = \sup_{f \in \cf} \cost(A,f,\varepsilon)
\]
For our trapezoidal rule algorithm, $A$
\[
\cost(A,\cf,\varepsilon) =  \cost(A,f,\varepsilon) = \left \lceil \sqrt{\frac{\rho}{8\varepsilon}} \right \rceil = \Order\left(\sqrt{ \frac{\rho}{\varepsilon}} \right) \qquad \forall f \in \cf = \{f :\norm[1]{f''} \le \rho \}  \medscoop
\]
This is the \alert{(up to a constant)} the smallest possible cost \uncover<2->{for integrands in $\cf = \{f :\norm[1]{f''} \le \rho \}  \medscoop$, no worse than Simpson's rule, Gauss rules, etc.}
\end{frame}

\finalthanksnote{Slides available on SpeakerDeck at
	\href{https://speakerdeck.com/fjhickernell/matrix-week-two-2018-june}
	{\nolinkurl{speakerdeck.com/fjhickernell/matrix-week-two-2018-june}}
}

\begin{frame}
\frametitle{Why the Trapezoidal Rule Is Optimal for $\cf = \{f :\norm[1]{f''} < \rho \}$}
\vspace{-3ex}
Let $A$ be \alert{any} algorithm satisfying
\[
\abs{\int_0^1 f(x) \, \dif x - A(f,\varepsilon)} \le \varepsilon \quad \forall f \in \cf= \{f :\norm[1]{f''} < \rho \}, \, \varepsilon > 0
\]

\vspace{-5ex}
\begin{tabular}{>{\raggedright}m{0.6\textwidth}@{\qquad}>{\centering}m{0.37\textwidth}}
	\begin{itemize}
		\item<2->Observe where $A(0,\varepsilon)$ gets function values

		\item<3->Construct $f_{\text{bump}} \in \cf$; note that $A(0,\varepsilon) = A(f_{\text{bump}},\varepsilon) = A(-f_{\text{bump}},\varepsilon)$

		\item<4->  For $A$ to succeed for $\pm f_{\text{bump}} $, which have different integrals, one must have 
		\[
		\cost(A,\cf,\varepsilon) =  \Order\left(\sqrt{ \frac{\rho}{\varepsilon}} \right)
		\]
		the same as the trapezoidal rule algorithm.

	\end{itemize}

	&
	\only<2>{\includegraphics[width=5cm]{ProgramsImages/ZeroFun.eps}}
	\only<3->{\includegraphics[width=5cm]{ProgramsImages/TwoQuadHump.eps}}
\end{tabular}

\uncover<5->{Why do your instructors tell you that other algorithms are \alert{better than the trapezoidal rule}?  \\  \uncover<6->{They are only better when you assume that the integrand has \alert{more smoothness}}}

\section{Adaptive Algorithms}

\begin{frame}
\frametitle{The Impracticality of Our Non-Adaptive Trapezoidal Rule Algorithm}
\vspace{-7ex}
\begin{tabular}{>{\raggedright}m{0.6\textwidth}@{\qquad}>{\centering}m{0.37\textwidth}}
\begin{gather*}
T_n(f)=\frac{1}{n}\left[\frac{f(0)}{2} + f\left(\frac{1}{n}\right) + \cdots + f\left(\frac{n-1}{n}\right) +  \frac{f(1)}{2} \right] \\
A(f,\varepsilon) = T_n(f) \quad \text{where }n = \left \lceil \sqrt{\frac{\rho}{8\varepsilon}} \, \right \rceil
\end{gather*}

\vspace{-2ex}
\begin{itemize}
	\item \alert{Solves} the integration problem for  $\cf = \{f :\norm[1]{f''} \le \rho \}$  \medscoop

	\item \alert{Costs} $\Order(\sqrt{\rho/\varepsilon})$ function values

>>>>

\newcommand{\sol}{S}
\newcommand{\app}{A}
\newcommand{\Sapp}{S_{\textup{app}}}
\newcommand{\LambdaStd}{\Lambda^{\textup{std}}}
\newcommand{\LambdaSer}{\Lambda^{\textup{ser}}}
\newcommand{\LambdaAll}{\Lambda^{\textup{all}}}
\DeclareMathOperator{\spann}{span}
%\DeclareMathOperator{\app}{app}

@@

\iffalse
Adaptive Approximation to Multivariate Linear Problems with Inputs Lying in a Cone

@@

\renewcommand{\OffTitleLength}{-10ex}
\setlength{\FJHThankYouMessageOffset}{-8ex}
\title{Adaptive Approximation for Multivariate Linear Problems  \\ with Inputs Lying in a Cone}
\author[]{Fred J. Hickernell}
\institute{Department of Applied Mathematics \\
	Center for Interdisciplinary Scientific Computation \\  Illinois Institute of Technology \\
	\href{mailto:hickernell@iit.edu}{\url{hickernell@iit.edu}} \quad
	\href{http://mypages.iit.edu/~hickernell}{\url{mypages.iit.edu/~hickernell}}}

@@

	\newcommand{\scoop}[1]{\parbox{#1}{\includegraphics[width=#1]{IceCreamScoop.eps}}\xspace}
	\newcommand{\smallscoop}{\scoop{1cm}}
	\newcommand{\medscoop}{\parbox{1.8cm}{\includegraphics[width=1.8cm]{IceCreamScoop.eps}}\xspace}
	\newcommand{\largescoop}{\scoop{3cm}}
	\newcommand{\ICcone}[1]{\parbox{#1}{\includegraphics[width=#1,angle=270]{MediumWaffleCone.eps}}\xspace}
	\newcommand{\medcone}{\ICcone{1.2cm}}
	\newcommand{\largercone}{\parbox{2.2cm}{\vspace*{-0.2cm}\includegraphics[width=1cm,angle=270]{MediumWaffleCone.eps}}\xspace}
	\newcommand{\largecone}{\ICcone{1.8cm}}
	\newcommand{\smallcone}{\parbox{0.65cm}{\includegraphics[width=0.3cm,angle=270]{MediumWaffleCone.eps}}\xspace}

@@

\begin{frame}{Multivariate Linear Problems}
\vspace{-3ex}
\begin{tabular}{p{0.47\textwidth}p{0.5\textwidth}}
Given $f \in \cf$ find $S(f) \in \cg$, where \\
$\sol: \cf \to \cg$ is linear, e.g., 
\begin{gather*}
    S(f) = \int_{\reals^d} f(\vx) \, \varrho(\vx) \, \dif \vx\\
    S(f) = f\\
    S(f) = \frac{\partial f}{\partial x_1} \\
    - \nabla^2 S(f) = f, \ \  S(f) = 0 \text{ on boundary}
\end{gather*}
&
\vspace{-5ex}
\alert{Successful algorithms}
\vspace{-3ex}
\begin{multline*}
    \ca(\cc) : = \{\app: \cc \times (0,\infty) \to \cg \text{ such that } \\
\norm[\cg]{\sol(f) - \app(f,\varepsilon) } \le \varepsilon \ \ \forall f \in \cc \subseteq \cf, \ \varepsilon > 0 \}
\end{multline*}

\vspace{-1ex}
where $\app(f,\varepsilon)$ depends on \alert{function values}, $\LambdaStd$, \alert{Fourier coefficients}, $\LambdaSer$, or \alert{linear functionals}, $\LambdaAll$, e.g.,

\vspace{-4ex}
\begin{gather*}
    \app(f,\varepsilon) = \Sapp(f,n) = \sum_{i=1}^n f(\vx_i) g_i, \quad g_i \in \cg \\
    \app(f,\varepsilon) = \Sapp(f,n) = \sum_{i=1}^n \hf_i g_i, \quad g_i \in \cg \\
    \app(f,\varepsilon) = \Sapp(f,n) = \sum_{i=1}^n L_i(f) g_i, \quad g_i \in \cg \\
\end{gather*}

\vspace{-5ex}
\hfill \hfill \alert{$n = ?$, $\cc$ is not a ball}

@@

\begin{tabular}{p{0.48\textwidth}p{0.49\textwidth}}
Given $f \in \cf$ find $S(f) \in \cg$, where \\
$\sol: \cf \to \cg$ is linear

\bigskip

\alert{Successful algorithms}
\vspace{-2ex}
\begin{multline*}
    \ca(\cc) : = \{\app: \cc \times (0,\infty) \to \cg \text{ such that } \\
\norm[\cg]{\sol(f) - \app(f,\varepsilon) } \le \varepsilon \ \ \forall f \in \cc \subseteq \cf, \ \varepsilon > 0 \}
\end{multline*}

\vspace{-1ex}
where $\app(f,\varepsilon)$ depends on \alert{function values}, $\LambdaStd$, \alert{Fourier coefficients}, $\LambdaSer$, or \alert{linear functionals}, $\LambdaAll$
&

\vspace{-9ex}
\uncover<2->{\alert{Solvability}\footfullcite{KunEtal19a}: $\ca(\cc) \ne \emptyset$

@@

\alert{Construction}: Identify $\app \in \ca(\cc)$}

@@

\uncover<3->{\alert{Cost}: $\cost(\app,f,\varepsilon) = $ \# of function data
\newline
 $\cost(\app,\cc,\varepsilon, \rho) = \max_{f \in \cc \cap \cb_{\rho}} \cost(\app,f,\varepsilon)$
 \newline
 $\cb_{\rho} := \{f \in \cf : \norm[\cf]{f} \le \rho \}$

@@

\alert{Complexity}\footfullcite{TraWasWoz88}: $\comp(\ca(\cc),\varepsilon,\rho)$ 
\newline \phantom{a} \hfill \hfill $= \min_{\app \in \ca(\cc)} \cost(\app,\cc,\varepsilon, \rho)$

@@

\alert{Optimality}:  \newline \phantom{a} \hfill \hfill $\cost(\app,\cc,\varepsilon, \rho) \le \comp(\ca(\cc),\alert{\omega} \varepsilon,\rho)$}

@@

\uncover<4->{\alert{Tractability}\footfullcite{NovWoz08a}: $\comp(\app,\cc,\varepsilon, \rho) \le C \rho^p\varepsilon^{-p} d^{q} $}

@@

\begin{frame}{Cones}
\vspace{-4ex}
\begin{tabular}{>{\centering}m{0.4\textwidth}@{\qquad}>{\centering}m{0.4\textwidth}}
     \largescoop \hspace{-3cm}\raisebox{-4ex}{\color{red}\fontsize{100}{120}\selectfont $\times$} & 
      \largecone \tabularnewline
      Ball $\cb_{\rho} := \{f \in \cf : \norm[\cf]{f} \le \rho \}$ &
      \hspace{2cm} (Non-Convex) Cone $\cc$
\end{tabular}

\begin{itemize}
    \item Assume set of inputs, $\cc \subseteq \cf$, is a \alert{cone}, not a ball

    \item \alert{Cone} means $f \in \cc \implies af \in \cc$

    \item \alert{Cones} are unbounded

    \item If we can bound the error for $f \in $ \alert{cone}, then we can typically also bound the error  for $af$

    \item Union of \alert{cones} is a \alert{cone}
\end{itemize}

\section{Solvability}
\begin{frame}{When Is $(S:\cc \subseteq \cf\to\cg,\Lambda)$ \emph{Solvable}?}

\vspace{-7ex}
\begin{multline*}
     \ca(\cc) : = \{\app: \cc \times (0,\infty) \to \cg \text{ such that } 
\norm[\cg]{\sol(f) - \app(f,\varepsilon) } \le \varepsilon \ \forall f \in \cc \subseteq \cf, \ \varepsilon > 0 \}
  \\
  \text{where $A(f,\varepsilon)$ depends on } \{L_i(f)\}_{i=1}^n \in \Lambda^n \subseteq \cf^{*n}
\end{multline*}

\alert{Definition} \quad $(S:\cc \subseteq \cf\to\cg,\Lambda)$ \alert{solvable} $\iff \ca(\cc) \ne \emptyset$\\[1ex]
\alert{Lemma}  \quad $f_1, f_2 \in \cc$  and $A(f_1, \varepsilon) = A(f_2,\varepsilon) \implies \norm[\cg]{S(f_1-f_2)} \le \varepsilon$\\[1ex]
\alert{Corollary} \label{ZeroCorollary} \quad $(S:\cc \subseteq \cf\to\cg,\Lambda)$ \alert{solvable} and $\exists f \in \cc,\  \varepsilon > 0$ with $A(f,\varepsilon) = A(0,\varepsilon)$  \\ \hfill \hfill $\implies S(f) = 0$\\[1ex]
\alert{Theorem} \label{VectorSpaceThm}\quad $(S:\cc \subseteq \cf\to\cg,\Lambda)$ \alert{solvable} and $\cc$ is a \alert{vector space} \\ \hfill \hfill 
$\iff S(f) = \sum_{i=1}^n L_i(f) g_i$ for some $\vL \in \Lambda^n, \vg \in \cg^n$  \hyperlink{VectorSpaceThmProof}{\beamergotobutton{Proof}}\\[1ex]
E.g. $\left(\int_{[0,1]^d} \cdot (\vx) \, \dif \vx: C^{1,\ldots,1}[0,1]^d\to\reals,\Lambda^{\textup{std/\alert{all}}}\right)$ is unsolvable/\alert{solvable}

\section{Series Spaces}

\begin{frame}{Spaces of Inputs and Outputs}

We focus on Banach spaces that look like

\vspace{-5ex}
\begin{align*}
    \cf &:= \left \{ f = \sum_{\vj \in \cn} \hf(\vj) u_{\vj} : \norm[\cf]{f} := \bignorm[r,\vlambda]{\hf} : = \norm[r]{\left(\frac{\bigabs{\hf(\vj)}}{\lambda_{\vj}} \right)_{\vj \in \cn}} \right \} \\
    \cg &: = \left \{ g = \sum_{\vj \in \cn} \hg(\vj) v_{\vj} : \norm[\cg]{g} := \bignorm[t]{\hg}\right \}, \qquad S(u_{\vj}) = v_{\vj}, \\
    \text{E.g., } & \text{ Function recovery: } S(u_{\vj}) = u_{\vj} = v_{\vj} \\
    & \text{ Integration: } S(u_{\vj}) = \int_{[0,1]^d} u_\vj (\vx) \, \dif \vx  = v_\vj\\
    \Sapp(f,n) &= \sum_{i=1}^n L_i(f) g_i = \sum_{\vj \in \cn} \hf(\vj) \Sapp(u_\vj,n)  \\
    \norm[\cg]{S(f) - A(f,\varepsilon)} & \le \sum_{\vj \in \cn} \bigabs{\hf(\vj)} \norm[\cg]{v_\vj - \Sapp(u_{\vj},n)}
\end{align*}

\begin{frame}
\frametitle{Example Bases}
\vspace{-3ex}
	\begin{tabular}{>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}}
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_0.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_3.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_4.png} 
	\tabularnewline[-7ex]
	Legendre
	\tabularnewline
	\tabularnewline
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2_3.png} 
\tabularnewline[0ex]
		\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_0.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_4.png} 
\tabularnewline[-7ex]
Chebyshev \tabularnewline
\tabularnewline
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2_3.png} 
	\end{tabular}
\end{frame}

\begin{frame}
\frametitle{Example Bases}
\vspace{-3ex}
	\begin{tabular}{>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}>{\centering}m{0.18\textwidth}}
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_0.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_3.png}  &
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_4.png} 
	\tabularnewline[-7ex]
	Legendre
	\tabularnewline
	\tabularnewline
		\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Legendre_Degree_2_3.png} 
\tabularnewline[0ex]
		\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_0.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_4.png} 
\tabularnewline[-7ex]
Chebyshev \tabularnewline
\tabularnewline
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_1.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_1_3.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2_2.png}  &
\includegraphics[width =0.18\textwidth]{FourierSampling/Chebyshev_Degree_2_3.png} 
	\end{tabular}
\end{frame}

\begin{frame}{Series Spaces}

\begin{align*}
    \cf &:= \left \{ f = \sum_{\vi \in \cn} \hf(\vi) u_{\vi} : \norm[\cf]{f} := \bignorm[r,\vlambda]{\hf} : = \norm[r]{\left(\frac{\bigabs{\hf(\vi)}}{\lambda_{\vi}} \right)_{\vi \in \cn}} \right \} \\
    \cg &: = \left \{ g = \sum_{\vi \in \cn} \hg(\vi) v_{\vi} : \norm[\cg]{g} := \bignorm[t]{\hg}\right \}, \qquad S(u_{\vi}) = v_{\vi}, \\
    \text{E.g., } & \text{ Function recovery: } S(u_{\vi}) = u_{\vi} = v_{\vi} \\
    & \text{ Integration: } S(u_{\vi}) = \int_{[0,1]^d} u_\vi (\vx) \, \dif \vx  = v_\vi\\
    \norm[\cg]{S(f) - A(f,\varepsilon)} & \le \sum_{\vi \in \cn} \bigabs{\hf(\vi)} \norm[\cg]{v_\vi - \Sapp(u_{\vi},n)}
\end{align*}
If $\gamma_{\vi_1} \ge \gamma_{\vi_2} \ge \cdots$, then $S_{\textup{app},n}(f) = \sum_{j=1}^n \hf(\vi_j) v_{\vi_j}$ is an optimal building block, and 
\begin{equation*}
    \norm[t]{S(f) - S_{\textup{app},n}(f)} = \Bignorm[r']{\left(\hf(\vi_j)\right)_{j=n+1}^{\infty}} \le \bignorm[r,\vgamma]{\hf} \Bignorm[qq'/(q-q')]{\left(\gamma_{\vi_j}\right)_{j=n+1}^{\infty}}
\end{equation*}

\begin{frame}{Cones that Make Sense for Series Spaces}
\begin{gather*}
    \cf := \left \{ f = \sum_{\vi \in \cn} \hf(\vi) u_{\vi} : \norm[\cf]{f} := \bignorm[q,\vgamma]{\hf} : = \norm[q]{\left(\frac{\bigabs{\hf(\vi)}}{\gamma_{\vi}} \right)_{\vi \in \cn}} \right \} \\
    \cg : = \left \{ g = \sum_{\vi \in \cn} \hg(\vi) v_{\vi} : \norm[\cg]{g} := \bignorm[q']{\hg}\right \}, \qquad S(u_{\vi}) = v_{\vi}, \\
    \text{If } \gamma_{\vi_1} \ge \gamma_{\vi_2} \ge \cdots, \text{ then } \hS_n(f) = \sum_{j=1}^n \hf(\vi_j) v_{\vi_j} \text{ is a good buidling block} \\
    \text{E.g., Function recovery: } S(u_{\vi}) = u_{\vi} = v_{\vi}
\end{gather*}

@@

\section{Bonus}
\begin{frame}[label = VectorSpaceThmProof]{Proof of Theorem for Solvability on a Vector Space}

Let the cone $\cc$ be a \alert{vector space} and let

\vspace{-3ex}
\begin{itemize}
    \item $A$ be a successful algorithm

    \item $\varepsilon > 0$ be any positive tolerance

    \item  $\{L_1, \ldots, L_{M}\} \subset \Lambda$ be the linear functionals used by $A(0,\varepsilon)$, and

    \item  $\{L_1, \ldots, L_m\}$ be a basis for $\spann(\{L_1, \ldots, L_{M}\})$

    \item $n  = \min(m,\dim(\cc))$

        \item $\{f_1, \ldots, f_n\} \subset \cc$ satisfy $L_i(f_j) = \delta_{i,j}$, $i =1, \ldots, n$, $j=1, \ldots, m$
\end{itemize}

For any $f \in \cc$, let $\tf = f- \sum_{i=1}^n L_i(f) f_i$, and note that $L_j(\tf) = 0$ for $j =1, \ldots, M$.  Thus, $A(\tf,\varepsilon) = A(0,\varepsilon)$, and so by the \hyperlink{ZeroCorollary}{\beamergotobutton{Corollary}},
\[
0 = S(\tf) = S(f) - \sum_{i=1}^n L_i(f) S(f_i), \qquad \text{which implies } S(f) = \sum_{i=1}^n L_i(f) S(f_i)
\]

\vspace{-3ex}
    \hyperlink{VectorSpaceThm}{\beamerreturnbutton{Back}}
\end{frame}

@@
